Pre-processing high frequency animal tracking data

### 2020-08-13 01_background.Rmd

@@

This paper refers extensively to the `R` package `atlastools` (Gupte 2020), which can be installed from an online repository.

@@

>>>>

---
output: html_document
editor_options: 
  chunk_output_type: console
---

# Getting started

This section covers:

1. Installing the `R` package `atlastools`,

2. Simulating some realistic looking movement data using the `R` package `smoove` from [@gurarie2017], and

3. Introducing positioning errors into the simulated movement data.

@@

This paper refers extensively to the `R` package `atlastools` (Gupte 2020), which can be installed from [Github.](https://github.com/pratikunterwegs/atlastools). 
See the repository for information on the build status.
The code chunk below shows how to install `atlastools`.

@@

## Simulating movement data

Here, we simulate some movement data using the `smoove` `R` package from [@gurarie2017].

```{r prep_libs_01_02}
# load smoove and datatable
library(smoove)
library(data.table)
```

```{r sim_smoove_data}
# define parameters
nu <- 2
tau <- 1
dt <- .01 
ucvm1 <- simulateUCVM(nu = nu, tau = tau, T.max = 100, dt = dt)

# get the data in data.table format
data <- as.data.table(ucvm1$XY)

# rescale coordinates
data[, c("x", "y") := (lapply(.SD, 
                          scales::rescale)),
     .SDcols = c("x", "y")]

# add a time column
data[, time := seq_len(nrow(data))]

# save simulated data
fwrite(data, "data/data_sim.csv")
```

Here, we write functions to add normally distributed errors to the position coordinates of the simulated data.

```{r fun_add_error}
# simple function to introduce normally distributed error
do_add_error <- function(v, std_dev = 0.5) {
  pos_error <- stats::rnorm(length(v), sd = std_dev)
  return(v + pos_error)
}
```

Example of introduced error.

```{r}
# load data
data <- fread("data/data_sim.csv")

# introduce errors to data
data[, `:=`(x_error = do_add_error(x, std_dev = 0.01),
            y_error = do_add_error(y, std_dev = 0.01))]

# plot real and introduced error
ggplot(d2)+

  geom_path(aes(x_error, y_error),
            col = "red", size = 0.2)+
  geom_path(aes(x, y))##+
  geom_path(data = data, aes(x, y),
            col = "blue")

atlastools::atl_compare_tracks(d2, data)
```
